---
title: '**First Part: COVID-19 in Spain**'
subtitle: '**Advanced Regression And Prediction**'
author: '*Roberto Jesús Alcaraz Molina*'
date: "09/05/2021"
output:
  pdf_document:
    latex_engine: xelatex
    number_section: true
    highlight: tango
    fig_caption: yes
    toc: true
    # toc_depth: 4
header-includes: 
- \usepackage{float}
- \usepackage{amsbsy}
- \usepackage{amsmath}
- \usepackage{graphicx}
- \usepackage{subfig}
- \usepackage{booktabs}
bibliography: references.bib
---

```{=tex}
\begin{center}
\includegraphics[width=3in]{logouc3m}
\end{center}
```
\newpage

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      echo = F, 
                      message = FALSE,
                      eval = T,
                      fig.pos="H", 
                      fig.align="center",
                      fig.width=11,
                      cache=FALSE, error = TRUE)
```

```{r, echo = F}
# Model packages
pacman::p_load(ranger)

# Packages
# devtools::install_github("stevenpawley/recipeselectors")
pacman::p_load(COVID19, tidymodels, rsample, recipes, parsnip, 
               yardstick, workflows, tune, patchwork, lubridate, recipeselectors,
               doParallel, tidyquant, finetune, ggpubr, MASS, modeltime, timetk,
               stacks, GGally)
library(tidyverse, attach.required = T)
# For ggplot
theme_set(theme_tq())

spain <- readRDS("../00_data/spain.RDS")
spain_new_data <- readRDS("../00_data/spain_new_data.RDS")
# spain_new_data <- covid19(country = "spain", start = "2021-04-11", end = "2021-05-02")
```




# INTRODUCTION

This project is the second and final project of the *Advanced Regression and Prediction* course. As we did in the first part, we are going to analyze and try predict the 
behavior of the coronavirus crisis in Spain, but this time, we are going to use
**machine learning** tools.

Even though for the first part of the project we took all days from 22nd of January 2020 until the 11th of March 2021, we will consider also all new observations until the
2nd of May, since the more data, the better will be our analysis.

This project will be divided in four parts: firstly, we are going to review the
data cleaning and preprocessing steps we did in the previous part and comment about
the results we had; then we will start the machine learning process, where we will
explain, tune and apply 6 differents machine learning models (K-Nearest Neighbors (KNN),
Support Vector Regression (SVR), Regression Trees, Random Forests, XGBoost and Neural
Network (NN)); after that, we will select the best model in addition to creating
an ensemble model with the best set of models; and finally we will give our conclusions.

To put us in context, we can remember the preprocessing we did to our data:

- First of all, we selected the target variables of our study, which are the deaths
and confirmed cases, and some other variables that could be interesting for our model.

- Secondly, we realized that we had some missing values in our two variables of 
interest, but they were easy to impute since they indicated that there were no cases.

- After that, we recategorized some of our factor variables because they were encoded
as integer, as well as removing and modifiying some levels that were not entirely true 
in Spain.

- Next, we fixed some errors that were related with the data collection, for instance, 
there were some days that had lower confirmed cases or deaths than the day before.

- Then, we aggregated the data to have **weekly data** in order to decrease the noise
that appear due to different human errors. For the numeric variables, we added all 
values, and for the categorical features, we took the mode, i.e., the most frequent
value in every week. Below, we can observe how our main variables of interest look like:

```{r, fig.cap="Weekly deaths and confirmed cases due to COVID-19 in Spain"}
p1 <- spain %>%
  plot_time_series(.date_var = date, 
                   deaths_week, 
                   .smooth = F, 
                   .interactive = F, .line_size = 1,
                   .title = "Weekly deaths")

p2 <- spain %>%
  plot_time_series(.date_var = date, 
                   confirmed_week, 
                   .smooth = F, 
                   .interactive = F, .line_size = 1,
                   .title = "Weekly confirmed cases")

ggarrange(p1, p2) + ggtitle("COVID-19 in Spain")
```

- Finally, we did some feature engineering, adding four new variables that took into 
account the time (month and year), and past cases (deaths and confirmed cases 3 weeks).
Therefore, the variables we finally considered for our model can be seen below. After all 
these steps, we are in position to start with the machine learning process.

| **Variable**                     | **Description**                                          |
|:---------------------------------|----------------------------------------------------------|
| `date`                           | Observation date (identifier).                           |
| `deaths_week`                    | Number of deaths per week.                               |
| `lag_3_deaths_week`              | Number of deaths per week three weeks ago.               |
| `confirmed_week`                 | Number of confirmed cases per week.                      |
| `lag_3_confirmed_week`           | Number of confirmed cases per week three weeks ago.      |
| `vaccines_week`                  | Number of doses administered (single dose) per week.     |
| `month`                          | Observation month .                                      |
| `year`                           | Observation year.                                        |
| `stay_home`                      | Indicates the measures of staying at home.               |
| `school_closing`                 | Indicates the measures in education.                     |
| `workplace_closing`              | Indicates the measures of the workplace.                 |
| `gatherings_restrictions`        | Indicates the measures of gatherings.                    |
| `internal_movement_restrictions` | Indicates the measures of the movements between regions. |
: Description of the variables

\bigskip
# MACHINE LEARNING

In the following chapter, some well-known machine learning tools will be applied.
We will explain the preprocessing steps that every algorithm needs, which are their main
hyperparameters and how they work in order to tune their hyperparameters and draw
comparison between the models.

First of all, since we have new data, we have to repeat the recursive feature
selection to select the most important features by means of a Random Forest. For the 
data set that will be used for deaths models, it removes the variables `stay_home` 
and `gathering_restrictions`, nonetheless, for the confirmed cases, it removes 
`workplace_closing` and again `gathering_restrictions`.

```{r}
rfe_model <- rand_forest() %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("regression")

# Predictive model for the deaths
set.seed(1234)
select_rec_deaths <- 
  recipe(deaths_week ~ ., data = spain) %>%
  step_rm(date) %>%
  step_select_vip(all_predictors(), outcome = "deaths_week", model = rfe_model, threshold = 0.2)

spain_deaths <- select_rec_deaths %>% prep() %>% juice()
spain_deaths <- spain_deaths %>%
  mutate(date = spain$date)

# saveRDS(spain_deaths, "../00_data/spain_deaths.RDS")

# Predictive model for the confirmed cases
select_rec_confirmed <- 
  recipe(confirmed_week ~ ., data = spain) %>%
  step_rm(date) %>%
  step_select_vip(all_predictors(), outcome = "confirmed_week", model = rfe_model, threshold = 0.2)


spain_confirmed <- select_rec_confirmed %>% prep() %>% juice()
spain_confirmed <- spain_confirmed%>%
  mutate(date = spain$date)

# saveRDS(spain_confirmed, "../00_data/spain_confirmed.RDS")
```


Next, we need to divide our training data into folds to tune the model parameters 
and to compare them. To do that, we will use again the function called `rolling_origin` 
from the `rsample` package (from `tidymodels`), but now there will be 36 weeks to train
in each fold, since as we saw in the previous part, 12 weeks were too little. Moreover,
the number of folds will increase up to 25 folds, so the parameters will be tuned in
a more efficient way.

After all these explanations, let's start with the first model: **K-Nearest Neighbors**.

```{r}
set.seed(123)
spain_rolling_deaths <- rolling_origin(
  spain_deaths, 
  initial = 36, # number of weeks used to train in each resample
  assess = 3, # number of weeks used to validate in each resample
  cumulative = F,
  skip = 0
  )

spain_rolling_confirmed <- rolling_origin(
  spain_confirmed, 
  initial = 36, # number of weeks used to train in each resample
  assess = 3, # number of weeks used to validate in each resample
  cumulative = F,
  skip = 0
  )
```


```{r}
grid_ctrl <- control_grid(save_pred = T, save_workflow = T)

basic_rec_deaths <- 
  recipe(deaths_week ~ ., data = spain_deaths) %>%
  step_rm(date)

basic_rec_confirmed <- 
  recipe(confirmed_week ~ ., data = spain_confirmed) %>%
  step_rm(date)
```



## K-Nearest Neighbors (K-NN)

The nearest neighbors algorithm **(k-NN)** is a supervised machine learning model
normally  used for both classification and regression. Firstly, it is important to
define a distance between the observations, usually the euclidean is the most popular.
Then, the distance between the observation to classify and all points in the sample are
calculated. After that, we have to select a number for $k$, which will be the number of
the $k$ closest observations from the point that we want to classify. Finally, the observation will be classified as the average of these others $k$ observations.

Regarding the preprocessing steps, there are not too many assumptions, but there are
some changes that are recommended. For instance, we will decorrelate predictors because
it improves the prediction accuracy even though it is not mandatory. Also, we will 
convert all categorical variables to dummy (note that most of the algorithms do it 
without any specification), we will normalize all predictors and finally, if necessary, 
we will remove those features with zero variance (those that only have one value).

With respect to the hyperparameter tuning, we have to focus on two main parameters:
neighbors (obviously), and the weight function (`weight_func`), which is the 
type of kernel function that weights the distance between samples (*rectangular*, 
*triangular*, *epanechnikov*, etc). Also, we will use the Euclidean distance to carry 
out these steps.


```{r}
# Deaths
kknn_deaths_rec <- 
  basic_rec_deaths %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# Confirmed
kknn_confirmed_rec <- 
  basic_rec_confirmed %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# KNN
kknn_model <- 
  nearest_neighbor(neighbors = tune(), weight_func = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("kknn")
```


## Support Vector Regression (SVR)

The support vector machine, or preferably, the support vector regression **(SVR)**
is an algorithm used for both classification and regression. In regression it fits
a function to create a tube with width $\varepsilon$, where the points inside that 
tube will be used to predict the new points.

For this algorithm, the preprocessing steps we need to do are the same that we did
for the k-NN algorithm: decorrelate predictors, create dummy variables, normalize 
the predictors and remove those with zero variance.

We will consider two algorithms from this family: the *polynomial SVR* and the 
*radial basis function SVR*. The difference between them is the kernel function used
to map the original data set into a higher dimensional space. Therefore, for the 
polynomial SVR we will have a parameter for the polynomial degree (`degree`) and another
one for the scaling factor (`scale_factor`) for the kernel; and for the rbf SVR we will 
have an hyperparameter $\sigma$ (`rbf_sigma`), which determines the reach of a single 
training instance. Also, there are another two common hyperparameters for both 
algorithms: the cost (`cost`) of predicting a  sample within or on the wrong side of the
margin, and the $\varepsilon$ (`margin`) we have said before.


```{r}
# Deaths
svm_deaths_rec <-
  basic_rec_deaths %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# Confirmed
svm_confirmed_rec <-
  basic_rec_confirmed %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# SVM
svm_poly_model <-
  svm_poly(cost = tune(), degree = tune(), scale_factor = tune(), margin = tune()) %>%
  set_engine('kernlab') %>%
  set_mode('regression')

svm_rbf_model <-
  svm_rbf(cost = tune(), rbf_sigma = tune(), margin = tune()) %>%
  set_engine('kernlab') %>%
  set_mode('regression')
```


## Regression Trees, Random Forest and eXtreme Gradient Boosting

A decision or regression tree **(RT)** is the result of asking an ordered sequences of
questions finishing in a prediction for the observation. The starting point of the
classification tree is called root node and consist of the entire training set in the
top of the tree. A node is a subset of the set of variables and it can be a parent node,
which is a node that splits into another two nodes if the condition is satisfied or not,
and a terminal node, which is a node that does not split and is assigned a value. This
algorithm has the advantage of interpretability or small bias, however, the variance
tend to be very large. Therefore, to reduce that variance, we combine some decision
trees, which leeds to Random Forest.

The Random Forest **(RF)** algorithm is an ensemble of regression trees. Firstly, we
start with $B$ boostrap samples drawn from the training set, which reduce variance, due
to aggregation, and bias, if the trees are fully grown. For each sample, a decision tree
is grown. Then, we predict every observation in each tree, and the final value for 
the observation will be the average of the results of each tree.

The third one, e**X**treme **G**radient **B**oosting or **XGBoost**, is a decision tree
ensemble machine learning algorithm that, unlike random forest, uses a boosting 
framework, i.e., instead of taking the average of multiple decision trees as the RF 
does, the models are built sequentially by minimizing the errors from previous models
while boosting the influence of high-performing models.

The principal advantages of the first two models is that they are scale invariant, i.e.,
it does not matter the scale of the features, so in these models we will not apply any 
preprocessing steps for the variables. Another advantage is that it is robust to 
irrelevant predictors, because if the variables is not very relevant, it is very 
possible that they will not use it to split the data. In contrast, for the XGBoost 
we should decorrelate predictors, create dummy variables and remove the zero variance
variables.

For the decision trees, we will have to tune three hyperparameters: a cost parameter
for the complexity of the tree (`Cp`), the maximum depth for the tree (`tree_depth`)
and the minimum number of observations in a node that are required for the node to
be split further (`min_n`). Secondly, for the Random Forest we will need to
tune the number of predictors that will be randomly sampled at each split when creating
decision trees (`mtry`), the number of trees contained in the ensemble (`trees`) and
again, `min_n`. Finally, for the XGBoost, we have the same hyperparameters than the
RF, the `tree_depth` from the decision tree and two additional hyperparameters: a number
for the rate at which the boosting algorithm adapts from iteration to iteration 
(`learn_rate`), and the reduction in the loss function required to split further 
(`loss_reduction`).


```{r}
# Deaths
dec_tree_deaths_rec <- 
  basic_rec_deaths

# Confirmed
dec_tree_confirmed_rec <- 
  basic_rec_confirmed %>% 
  step_corr(all_numeric_predictors(), threshold = 0.9)

# Decision tree
dec_tree_model <-
  decision_tree(tree_depth = tune(), min_n = tune(), cost_complexity = tune()) %>%
  set_engine('rpart') %>%
  set_mode('regression')
```


```{r}
# Deaths
ranger_deaths_rec <- 
  basic_rec_deaths

# Confirmed
ranger_confirmed_rec <- 
  basic_rec_confirmed

# RF
ranger_model <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("ranger") 
```

```{r}
xgboost_deaths_rec <- 
  basic_rec_deaths %>% 
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>% 
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

xgboost_confirmed_rec <- 
  basic_rec_confirmed %>% 
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>% 
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# Xgboost
xgboost_model <- 
  boost_tree(trees = tune(), min_n = tune(), tree_depth = tune(), learn_rate = tune(), 
    loss_reduction = tune(), sample_size = 1) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")
```

## Neural network

A multi-layer perceptron is a multivariate statistical technique that maps the
inputs variables nonlinearly with the output variable. Between the inputs and output
there is also hidden variables arranged in layers. The hidden and output layers are
called nodes, neurons or processing units. Therefore, to model our problem we will have 
$n$ inputs nodes, $X_1, \dots, X_n$ which are the variables, one or more layers of hidden nodes and an output node $Y$.

The preprocessing steps we need to follow are the same as in the majority of models:
decorrelate predictors, create dummy variables, normalize the data and remove zero-variance
features.

Regarding the hyperparameters, we have to tune four: the number of hidden units 
(`hidden_units`), a penalty for the amount of weight regularization to reduce overfitting
(`penalty`), the number of training iterations (`epoch`) and the activation function, which
could be linear, softmax, relu or elu (`activation`).



```{r}
# Deaths
mlp_deaths_rec <-
  basic_rec_deaths %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# Confirmed
mlp_confirmed_rec <-
  basic_rec_confirmed %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_corr(all_numeric_predictors(), threshold = 0.9) %>% # decorrelate predictors
  step_zv(all_predictors()) 

# NN
mlp_model <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune(), activation = tune()) %>%
  set_engine('nnet') %>%
  set_mode('regression')
```

## Model tuning and selection

Now that we have already explained all the models with their characteristics, it is 
time to tune their hyperparameters. To do that, since our data set is very small, 
we are going to consider a grid by default of size 100, which added to the 25 folds we have
to tune the parameters, results in fitting 2500 models for each algorithm. To compare and
tune them, we will select the one with lowest root mean square error (RMSE) and we will 
look also to the r-square (RSQ).

After fitting the first models, we will look at the model results to see possible 
errors and also, better parameter grids. We will give more appropiate values for 
the set of parameters based on the results obtained and we will again consider another
grid of size 100 around the grid we have specified. Therefore, we will obtain 25
different metrics for each set of parameters, so we will select the one with the lowest
average RMSE. These results can be seen below:



```{r}
# Deaths
workflow_tune_deaths <- workflow_set(
  
  preproc = list(
    kknn_deaths = kknn_deaths_rec,
    svm_poly_deaths = svm_deaths_rec,
    svm_rbf_deaths = svm_deaths_rec,
    dec_tree_deaths = dec_tree_deaths_rec,
    ranger_deaths = ranger_deaths_rec,
    xgboost_deaths = xgboost_deaths_rec,
    mlp_deaths = mlp_deaths_rec
  ),
  
  models = list(
    wf_1 = kknn_model,
    wf_2 = svm_poly_model,
    wf_3 = svm_rbf_model,
    wf_4 = dec_tree_model,
    wf_5 = ranger_model,
    wf_6 = xgboost_model,
    wf_7 = mlp_model
  ), 
  
  cross = F
)


workflow_tune_confirmed <- workflow_set(
  
  preproc = list(
    kknn_confirmed = kknn_confirmed_rec,
    svm_poly_confirmed = svm_confirmed_rec,
    svm_rbf_confirmed = svm_confirmed_rec,
    dec_tree_confirmed = dec_tree_confirmed_rec,
    ranger_confirmed = ranger_confirmed_rec,
    xgboost_confirmed = xgboost_confirmed_rec,
    mlp_confirmed = mlp_confirmed_rec
  ),
  
  models = list(
    wf_1 = kknn_model,
    wf_2 = svm_poly_model,
    wf_3 = svm_rbf_model,
    wf_4 = dec_tree_model,
    wf_5 = ranger_model,
    wf_6 = xgboost_model,
    wf_7 = mlp_model
  ), 
  
  cross = F
)
```


```{r, eval=F}
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

workflow_results_deaths <- workflow_tune_deaths %>%
  workflow_map(
    seed = 1503,
    resamples = spain_rolling_deaths,
    grid = 100,
    control = grid_ctrl,
    verbose = T
   )

stopCluster(cl)
```


```{r}
# workflow_results_deaths %>%
#   pull_workflow_set_result("mlp_deaths_wf_7") %>%
#   collect_metrics() %>%
#   filter(.metric == "rmse") %>%
#   arrange(mean) %>%
#   head(5)
# 
# workflow_results_deaths %>%
#   pull_workflow_set_result("mlp_deaths_wf_7") %>%
#   collect_metrics() %>%
#   filter(.metric == "rsq") %>%
#   arrange(desc(mean)) %>%
#   head(5)

kknn_param <- kknn_model %>%
  parameters() %>%
  update(neighbors = neighbors(c(5, 31)))

svm_poly_param <- svm_poly_model %>%
  parameters() %>%
  update(cost = cost(c(-1, 4)),
         scale_factor = scale_factor(c(-10, -3)))

svm_rbf_param <- svm_rbf_model %>%
  parameters() %>%
  update(cost = cost(c(-1,5)),
         rbf_sigma = rbf_sigma(c(-10, -2)))

dec_tree_param <- dec_tree_model %>%
  parameters() %>%
  update(cost_complexity = cost_complexity(c(-10, -5)),
         tree_depth = tree_depth(c(2, 30)), 
         min_n = min_n(c(2, 36))) 

ranger_param <- ranger_model %>%
  parameters() %>%
  update(mtry = mtry(c(2, 15)),
         trees = trees(c(200, 2000)),
         min_n = min_n(c(2, 36)))

xgboost_param <- xgboost_model %>%
  parameters() %>%
  update(trees = trees(c(10, 2000)),
         min_n = min_n(c(3, 36)),
         learn_rate = learn_rate(c(-10, -3)),
         loss_reduction = loss_reduction(c(-10, -3)))

mlp_param <- mlp_model %>%
  parameters() %>%
  update(hidden_units = hidden_units(c(2, 10)),
         penalty = penalty(c(-10, -2)),
         epochs = epochs(c(300, 1000)))

workflow_tune_deaths <- workflow_tune_deaths %>%
  option_add(param_info = kknn_param, 
             id = "kknn_deaths_wf_1") %>%
  option_add(param_info = svm_poly_param, 
             id = "svm_poly_deaths_wf_2") %>%
  option_add(param_info = svm_rbf_param, 
             id = "svm_rbf_deaths_wf_3") %>%
  option_add(param_info = dec_tree_param,
             id = "dec_tree_deaths_wf_4") %>%
  option_add(param_info = ranger_param,
             id = "ranger_deaths_wf_5") %>%
  option_add(param_info = xgboost_param,
             id = "xgboost_deaths_wf_6") %>%
  option_add(param_info = mlp_param, 
             id = "mlp_deaths_wf_7")
```

```{r, eval=F}
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

workflow_results_deaths <- workflow_tune_deaths %>%
  workflow_map(
    seed = 1503,
    resamples = spain_rolling_deaths,
    grid = 100,
    control = grid_ctrl,
    verbose = T
   )

stopCluster(cl)

saveRDS(workflow_results_deaths, "../02_results/workflow_results_deaths.RDS")
```

```{r}
# Now, the confirmed cases:
grid_ctrl <- control_grid(save_pred = T, save_workflow = T)

# KNN
kknn_model <- 
  nearest_neighbor(neighbors = tune(), weight_func = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("kknn") 

# SVM
svm_poly_model <-
  svm_poly(cost = tune(), degree = tune(), scale_factor = tune(), margin = tune()) %>%
  set_engine('kernlab') %>%
  set_mode('regression')

svm_rbf_model <-
  svm_rbf(cost = tune(), rbf_sigma = tune(), margin = tune()) %>%
  set_engine('kernlab') %>%
  set_mode('regression')

# Decision tree
dec_tree_model <-
  decision_tree(tree_depth = tune(), min_n = tune(), cost_complexity = tune()) %>%
  set_engine('rpart') %>%
  set_mode('regression')

# RF
ranger_model <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("ranger") 

# Xgboost
xgboost_model <- 
  boost_tree(trees = tune(), min_n = tune(), tree_depth = tune(), learn_rate = tune(), 
    loss_reduction = tune(), sample_size = 1) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")

# NN
mlp_model <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune(), activation = tune()) %>%
  set_engine('nnet') %>%
  set_mode('regression')
```


```{r, eval=F}
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

workflow_results_confirmed <- workflow_tune_confirmed %>%
  workflow_map(
    seed = 1503,
    resamples = spain_rolling_confirmed,
    grid = 100,
    control = grid_ctrl,
    verbose = T
   )

stopCluster(cl)
```

```{r}
# workflow_results_confirmed %>%
#   pull_workflow_set_result("dec_tree_confirmed_wf_4") %>%
#   pull(.notes)
# 
# workflow_results_confirmed %>%
#   pull_workflow_set_result("dec_tree_confirmed_wf_4") %>%
#   collect_metrics() %>%
#   filter(.metric == "rmse") %>%
#   arrange(mean) %>%
#   head(5)
# 
# workflow_results_confirmed %>%
#   pull_workflow_set_result("mlp_confirmed_wf_7") %>%
#   collect_metrics() %>%
#   filter(.metric == "rsq") %>%
#   arrange(desc(mean)) %>%
#   head(5)

kknn_param <- kknn_model %>%
  parameters() %>%
  update(neighbors = neighbors(c(5, 31)))

svm_poly_param <- svm_poly_model %>%
  parameters() %>%
  update(cost = cost(c(-2, 5)),
         scale_factor = scale_factor(c(-10, -3)))

svm_rbf_param <- svm_rbf_model %>%
  parameters() %>%
  update(cost = cost(c(-2,5)),
         rbf_sigma = rbf_sigma(c(-10, -2)))

dec_tree_param <- dec_tree_model %>%
  parameters() %>%
  update(cost_complexity = cost_complexity(c(-10, -1)),
         tree_depth = tree_depth(c(2, 20)), 
         min_n = min_n(c(2, 30))) 

ranger_param <- ranger_model %>%
  parameters() %>%
  update(mtry = mtry(c(2, 15)),
         trees = trees(c(200, 2000)),
         min_n = min_n(c(2, 36)))

xgboost_param <- xgboost_model %>%
  parameters() %>%
  update(trees = trees(c(10, 2000)),
         min_n = min_n(c(3, 36)),
         learn_rate = learn_rate(c(-10, -2)),
         loss_reduction = loss_reduction(c(-10, -3)))

mlp_param <- mlp_model %>%
  parameters() %>%
  update(hidden_units = hidden_units(c(2, 9)),
         penalty = penalty(c(-10, -2)),
         epochs = epochs(c(200, 1000)))

workflow_tune_confirmed <- workflow_tune_confirmed %>%
  option_add(param_info = kknn_param, 
             id = "kknn_confirmed_wf_1") %>%
  option_add(param_info = svm_poly_param, 
             id = "svm_poly_confirmed_wf_2") %>%
  option_add(param_info = svm_rbf_param, 
             id = "svm_rbf_confirmed_wf_3") %>%
  option_add(param_info = dec_tree_param,
             id = "dec_tree_confirmed_wf_4") %>%
  option_add(param_info = ranger_param,
             id = "ranger_confirmed_wf_5") %>%
  option_add(param_info = xgboost_param,
             id = "xgboost_confirmed_wf_6") %>%
  option_add(param_info = mlp_param, 
             id = "mlp_confirmed_wf_7")
```

```{r, eval=FALSE}
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

workflow_results_confirmed <- workflow_tune_confirmed %>%
  workflow_map(
    seed = 1503,
    resamples = spain_rolling_confirmed,
    grid = 100,
    control = grid_ctrl,
    verbose = T
   )


stopCluster(cl)

saveRDS(workflow_results_confirmed, "../02_results/workflow_results_confirmed.RDS")
# workflow_results_confirmed <- readRDS("../02_results/workflow_results_confirmed.RDS")
```

```{r}
workflow_results_deaths <- readRDS("../02_results/workflow_results_deaths.RDS")
workflow_results_confirmed <- readRDS("../02_results/workflow_results_confirmed.RDS")
```

```{r, fig.cap="Confidence interval for the metrics of the best set of hyperparameters of each model for the deaths"}
autoplot(workflow_results_deaths, select_best = T) +
  ggtitle("Results for the deaths model") +
  scale_color_tq() +
  theme_tq()
```

In the graphic above, we can observe the metrics obtained from the best set of 
parameters and their confidence interval. For the deaths model, the model that has
outperformed the rest is the **radial basis function support vector machines**, 
which has around 130 RMSE and almost 0.7 RSQ on average. The set of hyperparameters
can be seen below:

|  **Hyperparameters** |   `rbf_svm`    |
|----------------------|:--------------:|
|`cost`                |     18.2275    |
|`rbf_sigma`           |     0.00155    |
|`margin`              |     0.10298    |
:Final set of parameters for the radial basis function SVM

```{r}
#workflow_results_deaths %>%
#  pull_workflow_set_result("svm_rbf_deaths_wf_3") %>% 
#  select_best()
```



On the other hand, as we can see below, we have the results for the confirmed cases model. In 
this case, we can observe that the best model seem to be the decision tree, however, it does 
not have any value for RSQ. It occurs when the model predicts a single value for all samples, 
so it is not what we want. Therefore, we will select the second best in RMSE, which is also 
the best in RSQ: the **XGBoost**.

```{r, fig.cap="Confidence interval for the metrics of the best set of hyperparameters of each model for the confirmed cases"}
autoplot(workflow_results_confirmed, select_best = T) +
  ggtitle("Results for the confirmed cases model") +
  scale_color_tq() +
  theme_tq()
```


Its best set of hyperparameters is:

|  **Hyperparameters** |   `xgboost`    |
|----------------------|:--------------:|
|`trees`               |     $1279$     |
|`min_n`               |     $9$        |
|`tree_depth`          |     $6$        |
|`learn_rate`          |     $0.0008$   |
|`loss_reduction`      |     $2.11e-09$ |
:Final set of parameters for the XGBoost

```{r}
# Deaths
best_results_deaths <- 
   workflow_results_deaths %>% 
   pull_workflow_set_result("svm_rbf_deaths_wf_3") %>% 
   select_best(metric = "rmse")

best_workflow_deaths <- 
   workflow_results_deaths %>% 
   pull_workflow("svm_rbf_deaths_wf_3") %>% 
   finalize_workflow(best_results_deaths)

# Confirmed
best_results_confirmed <- 
   workflow_results_confirmed %>% 
   pull_workflow_set_result("xgboost_confirmed_wf_6") %>% 
   select_best(metric = "rmse")

best_workflow_confirmed <- 
   workflow_results_confirmed %>% 
   pull_workflow("xgboost_confirmed_wf_6") %>% 
   finalize_workflow(best_results_confirmed)

best_fit_deaths <- best_workflow_deaths %>%
  fit(spain)

best_fit_confirmed <- best_workflow_confirmed %>%
  fit(spain)

# saveRDS(best_fit_deaths, "../02_results/best_fit_deaths.RDS")
# saveRDS(best_fit_confirmed, "../02_results/best_fit_confirmed.RDS")
```


In the following section, in order to improve our models, we are going to make an 
stacking ensemble with the best set of models we have fitted before.

```{r}
# workflow_results_deaths %>%
#   pull_workflow_set_result("dec_tree_deaths_wf_4") %>%
#   collect_metrics() %>%
#   filter(.metric == "rmse") %>%
#   arrange(mean) %>%
#   head(5)
# 
# workflow_results_deaths %>%
#   pull_workflow_set_result("dec_tree_deaths_wf_4") %>%
#   collect_metrics() %>%
#   filter(.metric == "rsq") %>%
#   arrange(desc(mean)) %>%
#   head(5)
```

\bigskip
# STACKING ENSEMBLE

In this section, we are going to combine those models we have created in order to
make another one with better predictive power. All this process can be done thanks
to the R package `stacks` from `tidymodels`, which have the necessary tools to carry
it out. 

Basically, we start by creating a data set with the outcome in the validation set (in
our case, every 3 weeks validation sets) and all the predictions from each candidate
member. Once we have our data set ready, we will apply a model to select the best set 
of candidates. This process is usually called *metalearning*.

Then, since the outputs of each member can be highly correlated, we should consider
a model that skips some of the methods. Therefore, the models that are used in these
steps are the ones we used in the previous project: ridge regression, lasso or elastic
net, depending on the degree of regularization we want. After fitting one of these
models, we get a coefficient for each member, which is called *weight*, that determines
which models will be taken as final members of the ensemble. After we have the weights 
for every member of the ensemble, we can fit the model in all the data set and predict
on new data. 

For our deaths model, we add all candidate members to the stack object, having finally
619 candidate members from the 7 model definitions. We do not have the 700 candidates
(one for each set of hyperparameters for each model) because some models configurations
of the k-NN and decision tree had errors. After that, we fitted a elastic net model,
and below we can see their results:

```{r}
deaths_stack_data <- stacks() %>%
  add_candidates(workflow_results_deaths)
```

```{r, eval=FALSE}
set.seed(123)
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

deaths_stack_model <- 
  deaths_stack_data %>%
  blend_predictions(
    penalty = 10^(-6:-1),
    mixture = seq(0, 1, 0.1),
    metric = metric_set(rmse, rsq)
    )

deaths_stack_model_fit <-
  deaths_stack_data %>%
  blend_predictions(
    penalty = 10^(-6:-1),
    mixture = seq(0, 1, 0.1),
    metric = metric_set(rmse, rsq)) %>%
  fit_members()

stopCluster(cl)

saveRDS(deaths_stack_model, "../02_results/deaths_stack_model.RDS")
saveRDS(deaths_stack_model_fit, "../02_results/deaths_stack_model_fit.RDS")
```

```{r}
deaths_stack_model <- readRDS("../02_results/deaths_stack_model.RDS")
deaths_stack_model_fit <- readRDS("../02_results/deaths_stack_model_fit.RDS")
```

\bigskip
```{r, echo=T, eval=F}
-- A stacked ensemble model -------------------------------------

Out of 619 possible candidate members, the ensemble retained 10.
Penalty: 0.1.
Mixture: 1.

The 10 highest weighted members are:
# A tibble: 10 x 3
   member                    type             weight
   <chr>                     <chr>             <dbl>
 1 xgboost_deaths_wf_6_1_046 boost_tree 187867.     
 2 svm_rbf_deaths_wf_3_1_025 svm_rbf         0.492  
 3 mlp_deaths_wf_7_1_050     mlp             0.206  
 4 mlp_deaths_wf_7_1_056     mlp             0.155  
 5 mlp_deaths_wf_7_1_033     mlp             0.141  
 6 mlp_deaths_wf_7_1_040     mlp             0.132  
 7 mlp_deaths_wf_7_1_004     mlp             0.0824 
 8 mlp_deaths_wf_7_1_045     mlp             0.0806 
 9 mlp_deaths_wf_7_1_048     mlp             0.0184 
10 mlp_deaths_wf_7_1_075     mlp             0.00210
```
\bigskip

It finally select ten models: XGBoost, rfb SVM and 8 multilayer perceptrons. As we can 
observe, for the first model, it gives a very large weight. Below we can see why: it predicts very small numbers, so it needs a larger weight.

```{r}
t <- workflow_results_deaths %>%
  pull_workflow_set_result("xgboost_deaths_wf_6") %>%
  collect_predictions() %>% 
  filter(.config == "Preprocessor1_Model046") %>% 
  dplyr::select(.pred, deaths_week) %>% 
  head(5)

knitr::kable(t, caption = "Predictions from the XGBoost model")
```

Also, we need to ensure that the predictions are not very correlated. To do that, we can
observe a correlation plot between the ten selected members. There are a couple of
them that have a high correlation (>0.8) but the majority are fine.

```{r, fig.cap="Correlation plot between the selected members"}
deaths_stack_model_fit$data_stack %>%
  rename(xgboost_deaths_046 = xgboost_deaths_wf_6_1_046, 
         svm_rbf_deaths_025 = svm_rbf_deaths_wf_3_1_025, 
         mlp_deaths_050 = mlp_deaths_wf_7_1_050, 
         mlp_deaths_056 = mlp_deaths_wf_7_1_056,
         mlp_deaths_033 = mlp_deaths_wf_7_1_033, 
         mlp_deaths_040 = mlp_deaths_wf_7_1_040,
         mlp_deaths_004 = mlp_deaths_wf_7_1_004, 
         mlp_deaths_045 = mlp_deaths_wf_7_1_045, 
         mlp_deaths_048 = mlp_deaths_wf_7_1_048,
         mlp_deaths_075 = mlp_deaths_wf_7_1_075) %>%
  dplyr::select(xgboost_deaths_046, svm_rbf_deaths_025, mlp_deaths_050, mlp_deaths_056,
                mlp_deaths_033, mlp_deaths_040, mlp_deaths_004, mlp_deaths_045,
                mlp_deaths_048, mlp_deaths_075) %>% 
  ggcorr(label = T, digits = 2, label_alpha = T, label_round = 2)
```


```{r}
# spain_deaths %>%
#   bind_cols(predict(deaths_stack_model_fit, .)) %>%
#   dplyr::select(deaths_week, .pred) %>%
#   rmse(truth = deaths_week, estimate = .pred)
# 
# spain_deaths %>%
#   bind_cols(predict(deaths_stack_model_fit, .)) %>%
#   dplyr::select(deaths_week, .pred) %>%
#   rsq(truth = deaths_week, estimate = .pred)
# 
# spain_new_data %>%
#   bind_cols(predict(deaths_stack_model_fit, .)) %>%
#   dplyr::select(deaths_week, .pred) %>%
#   rsq(truth = deaths_week, estimate = .pred)
```

On the other hand, let's see the ensemble model for the confirmed cases. Below, 
we can observe that it only chooses 4: a rbf SVM, a multilayer perceptron and 
two decision trees, giving the largest weight to the SVM model. Also, as we can 
observe in the correlation plot, the two decision tree models are very correlated
but since the rest are not, and the weights are very low, it is not a big problem.

```{r}
confirmed_stack_data <- stacks() %>%
  add_candidates(workflow_results_confirmed)
```

```{r, eval=FALSE}
set.seed(123)
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

confirmed_stack_model <- 
  confirmed_stack_data %>%
  blend_predictions(
    penalty = 10^(-6:-1),
    mixture = seq(0, 1, 0.1),
    metric = metric_set(rmse, rsq))

confirmed_stack_model_fit <-
  confirmed_stack_data %>%
  blend_predictions(
    penalty = 10^(-6:-1),
    mixture = seq(0, 1, 0.1),
    metric = metric_set(rmse, rsq)) %>%
  fit_members()

stopCluster(cl)
saveRDS(confirmed_stack_model, "../02_results/confirmed_stack_model.RDS")
saveRDS(confirmed_stack_model_fit, "../02_results/confirmed_stack_model_fit.RDS")
```


```{r}
confirmed_stack_model <- readRDS("../02_results/confirmed_stack_model.RDS")
confirmed_stack_model_fit <- readRDS("../02_results/confirmed_stack_model_fit.RDS")
```


```{r, fig.cap="Weights for the selected members and correlation plot between them"}
p1 <- autoplot(confirmed_stack_model, type = "weights")
p2 <- confirmed_stack_model_fit$data_stack %>%
  rename(svm_rbf_025 = svm_rbf_confirmed_wf_3_1_025, 
         mlp_015 = mlp_confirmed_wf_7_1_015, 
         dec_tree_058 = dec_tree_confirmed_wf_4_1_058, 
         dec_tree_002 = dec_tree_confirmed_wf_4_1_002
         ) %>%
  dplyr::select(svm_rbf_025, mlp_015, dec_tree_058, dec_tree_002) %>% 
  ggcorr(label = T, digits = 2, label_alpha = T, label_round = 2) +
  ggtitle("Correlation plot")


ggarrange(p1, p2)
```

Once we have all these models fitted and selected, we are going how they predict
in the training and testing sets, and we will give our conclusions.


```{r}
# spain_confirmed %>%
#   bind_cols(predict(confirmed_stack_model_fit, ., members = T)) %>%
#   dplyr::select(confirmed_week, .pred) %>%
#   rmse(truth = confirmed_week, estimate = .pred)
# 
# spain_confirmed %>%
#   bind_cols(predict(confirmed_stack_model_fit, .)) %>%
#   dplyr::select(confirmed_week, .pred) %>%
#   rsq(truth = confirmed_week, estimate = .pred)
# 
# spain_new_data %>%
#   bind_cols(predict(confirmed_stack_model_fit, ., members = T)) %>% # dplyr::select(confirmed_week, .pred)
```


\bigskip
# PREDICTIONS AND CONCLUSIONS

As we saw above, the best model for the deaths was the radial basis function SVM,
and for the confirmed cases, the XGBoost model. First of all, we are going to see
both model predictions in the training and testing sets and then, we will see
the behavior of the ensemble models.

\bigskip
```{r, fig.cap="Predictions in the training set from the radial basis function SVM (deaths) and XGBoost (confirmed)"}
pred_train_deaths <- best_fit_deaths %>%
  predict(new_data = spain)

p1 <- spain %>%
  mutate(predictions = pred_train_deaths$.pred) %>%
  mutate(predictions = ifelse(predictions <0, 0, predictions)) %>% 
  dplyr::select(date, deaths_week, predictions) %>%
  pivot_longer(cols = c("deaths_week", "predictions"), values_to = "value", names_to = "Variable") %>%
  ggplot(aes(x = date, y = value)) +
  geom_line(aes(color = Variable), size = 1) +
  scale_color_manual(values = c("darkred", "steelblue")) +
  labs(
    y = "Deaths per week",
    x = "Date"
  ) 

pred_train_confirmed <- best_fit_confirmed %>%
  predict(new_data = spain)

p2 <- spain %>%
  mutate(predictions = pred_train_confirmed$.pred) %>%
  mutate(predictions = ifelse(predictions < 0, 0, predictions)) %>% 
  dplyr::select(date, confirmed_week, predictions) %>%
  pivot_longer(cols = c("confirmed_week", "predictions"), values_to = "value", names_to = "Variable") %>%
  ggplot(aes(x = date, y = value)) +
  geom_line(aes(color = Variable), size = 1) +
  scale_color_manual(values = c("darkred", "steelblue")) +
  labs(
    y = "Confirmed cases per week",
    x = "Date"
  ) 

ggarrange(p1, p2) +
  ggtitle("Results in the training set of both models") +
  theme(plot.title = element_text(hjust = 0.5, size = 20))
```

\bigskip
As we can observe, both models work in a good way but they are missing some peaks.
For the deaths model, it does not get the peaks in April 2020 and either in April 2021.
In the confirmed cases model it happens practically the same, in January 2021 it
predicts a much lower value than it was. Now, let's see the predictions in the testing
set:

\bigskip
```{r}
pred_test_deaths <- best_fit_deaths %>%
  predict(new_data = spain_new_data)

pred_test_confirmed <- best_fit_confirmed %>%
  predict(new_data = spain_new_data)

results1 <- spain_new_data %>%
  mutate(
    predicted_deaths = pred_test_deaths$.pred,
    predicted_confirmed = pred_test_confirmed$.pred
  ) %>%
  dplyr::select(date, deaths_week, predicted_deaths, confirmed_week, predicted_confirmed) %>%
  summarise(
    Date = date,
    Deaths = deaths_week, 
    `Pred. Deaths` = round(predicted_deaths, digits = 0),
    `Dif. Deaths` = Deaths - `Pred. Deaths`,
    Confirmed = confirmed_week, 
    `Pred. Confirmed` = round(predicted_confirmed, digits = 0),
    `Dif. Confirmed` = Confirmed - `Pred. Confirmed`
  )

knitr::kable(results1, caption = "Results in the testing set")
# saveRDS(results1, "../02_results/results1.RDS")
```

The results in the testing sets are ver good. For the deaths model, the difference 
in the first week is very large but for the next two, the differences are very similar.
Moreover, for the confirmed model, the second week is almost perfectly predicted,
and the other differences are not very large, which is very good. Now, let's see
the results for the ensemble models:



```{r, fig.cap="Results in the training set for the ensemble models"}
pred_train_deaths <- deaths_stack_model_fit %>%
  predict(new_data = spain)

p1 <- spain %>%
  mutate(predictions = pred_train_deaths$.pred) %>%
  mutate(predictions = ifelse(predictions <0, 0, predictions)) %>% 
  dplyr::select(date, deaths_week, predictions) %>%
  pivot_longer(cols = c("deaths_week", "predictions"), values_to = "value", names_to = "Variable") %>%
  ggplot(aes(x = date, y = value)) +
  geom_line(aes(color = Variable), size = 1) +
  scale_color_manual(values = c("darkred", "steelblue")) +
  labs(
    y = "Deaths per week",
    x = "Date"
  ) 

pred_train_confirmed <- confirmed_stack_model_fit %>%
  predict(new_data = spain)

p2 <- spain %>%
  mutate(predictions = pred_train_confirmed$.pred) %>%
  mutate(predictions = ifelse(predictions < 0, 0, predictions)) %>% 
  dplyr::select(date, confirmed_week, predictions) %>%
  pivot_longer(cols = c("confirmed_week", "predictions"), values_to = "value", names_to = "Variable") %>%
  ggplot(aes(x = date, y = value)) +
  geom_line(aes(color = Variable), size = 1) +
  scale_color_manual(values = c("darkred", "steelblue")) +
  labs(
    y = "Confirmed cases per week",
    x = "Date"
  ) 

ggarrange(p1, p2) +
  ggtitle("Results in the training set of both ensemble models") +
  theme(plot.title = element_text(hjust = 0.5, size = 20))
```

Now, it can be seen that the predictions fits way better to the real values, reaching 
the peaks that, with the previous models, we were not able to get. Let's see the results
in the testing set.

```{r}
pred_test_deaths <- deaths_stack_model_fit %>%
  predict(new_data = spain_new_data)

pred_test_confirmed <- confirmed_stack_model_fit %>%
  predict(new_data = spain_new_data)

results2 <- spain_new_data %>%
  mutate(
    predicted_deaths = ifelse(pred_test_deaths$.pred < 0, 0, pred_test_deaths$.pred),
    predicted_confirmed = ifelse(pred_test_confirmed$.pred < 0, 0, pred_test_confirmed$.pred)
  ) %>%
  dplyr::select(date, deaths_week, predicted_deaths, confirmed_week, predicted_confirmed) %>%
  summarise(
    Date = date,
    Deaths = deaths_week, 
    `Pred. Deaths` = round(predicted_deaths, digits = 0),
    `Dif. Deaths` = Deaths - `Pred. Deaths`,
    Confirmed = confirmed_week, 
    `Pred. Confirmed` = round(predicted_confirmed, digits = 0),
    `Dif. Confirmed` = Confirmed - `Pred. Confirmed`
  )

knitr::kable(results2, caption = "Results in the testing set for the ensemble model")
# saveRDS(results2, "../02_results/results2.RDS")
```

Now, the results are much worse, and it may happen due to overfitting. Due to the 
small quantity of data we have, an ensemble does not seem to be the best option.
It learns in a very optimal way the training set but it has huge errors in the 
testing set, so at the end, it is not an useful model to predict.

As a conclusion, we have to admit that the results are very good taking into account
the size of the data set, which only has 63 weeks and the quickness that the situation
is changing. For instance, everybody knows that thanks to the vaccination process
to old people, deaths will decrease drastically and also since the people are getting
used to this pandemic time, the confirmed cases also goes down. However, those variables
are not in our data set and could improve much more the predictions, but again, 
with only 63 weeks it is very difficult to get prefect predictions. 

\newpage
# REFERENCES

- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source
  Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
  
- Kuhn et al., (2020). Tidymodels: a collection of packages for modeling and
  machine learning using tidyverse principles. https://www.tidymodels.org

- Guidotti, E., Ardia, D., (2020), "COVID-19 Data Hub", Journal of Open Source Software 5(51):2376.

- Steven Pawley (2021). recipeselectors: Extra Recipes Steps for
  Supervised Feature Selection. R package version 0.0.1.
  https://github.com/stevenpawley/recipeselectors

- Kuhn, Max, and Kjell Johnson. 2019. “Feature Selection Overview.” In, 227–40. Chapman; Hall/CRC.

